Reproducible Research: Peer Assessment 1
=======================================

# Loading and preprocessing the data
1. Load the data
```{r}
unzip("repdata-data-activity.zip")
activity <- read.csv("activity.csv")
```
2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
activityfilterd <- activity[!is.na(activity$steps), ]
```

# What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
stepsperday <- aggregate(steps ~ date, activityfilterd, sum)
```
2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day.
```{r}
barplot(stepsperday$steps)
hist(stepsperday$steps)
```
3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(stepsperday$steps)
median(stepsperday$steps)
```

# What is the average daily activity pattern?
1. Make a time series plot (i.e. ```type = "l"```) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
stepsinday <- aggregate(steps ~ interval, activityfilterd, mean)
with(stepsinday, plot(interval, steps, type = "l"))
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
stepsinday[which.max(stepsinday$steps),]$interval
```

# Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with ```N/A``` s)
```{r}
sum(apply(is.na(activity), MARGIN=c(1), max))
```
It turns out this is the same as total number of missing values in ```steps``` column
```{r}
sum(is.na(activity$steps))
```
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc. (I'm using the mean for that 5-minute interval)
```{r}
nas <- is.na(activity$steps)
filledsteps <- stepsinday[match(activity[nas, ]$interval, stepsinday$interval), ]$steps
```
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityfilled <- activity
activityfilled[nas,]$steps <- filledsteps
```
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
stepsperdayfilled <- aggregate(steps ~ date, activityfilled, sum)
hist(stepsperdayfilled$steps)
mean(stepsperdayfilled$steps)
median(stepsperdayfilled$steps)
```
Do these values differ from the estimates from the first part of the assignment? 
```{r, echo = FALSE}
print("The mean value remains the same, but median value changed.")
```
What is the impact of inputing missing data on the estimates of the total daily number of steps?
```{r, echo = FALSE}
print("Total number of data points increased. histogram has more counts in the center bar.")
```
# Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels ¨C ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day.
```{r}
weekdayorweekend <- sample("weekday",nrow(activityfilled), replace = TRUE)
weekdayorweekend[grepl("Sat|Sun", weekdays(as.Date(activityfilled$date), abbr = TRUE))] <- "weekend"
weekdayorweekend <- factor(weekdayorweekend)
```
2. Make a panel plot containing a time series plot (i.e. ```type = "l"```) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
stepsinweekday <- aggregate(steps ~ interval, activityfilled[weekdayorweekend == "weekday", ], mean)
stepsinweekend <- aggregate(steps ~ interval, activityfilled[weekdayorweekend == "weekend", ], mean)
stepsinweekday$date <- "weekday"
stepsinweekend$date <- "weekend"
intervaldata <- rbind(stepsinweekday, stepsinweekend)
library(lattice)
xyplot(steps ~ interval | date, intervaldata, type = "l", layout = c(1,2))
```